---
title: "ridership"
author: "Christine Iyer"
date: "February 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(forecast)
```

#Read in Data
```{r}
ridership <- read.csv("Amtrak.csv")
head(ridership)
```

#Convert factor to date and plot the data
```{r}
str(ridership)
ridership$Month <- as.character(as.Date(ridership$Month, format = "%m/%d/%Y"))
head(ridership)
```

```{r}
ridership.ts <- ts(ridership$Ridership_in_thousands, start = c(1991, 1), end = 2004, 3, freq = 12)
plot(ridership.ts, xlab = "Time", ylab = "Ridership", ylim = c(1300, 2200), bty = "l")



```

#Overlaid polynomial trend line
```{r}
ridership.lm <- tslm(ridership.ts ~ trend + I(trend^2))
par(mfrow = c(2,1))
plot(ridership.ts, xlab = "Time", ylab = "Ridership", ylim = c(1300, 2300), bty = "l")
lines(ridership.lm$fitted, lwd = 2)
ridership.ts.zoom <- window(ridership.ts, start = c(1997, 1), end = c(2000, 12))
#zoom

```

#Zoom into 4 years of data
```{r}
plot(ridership.ts.zoom, xlab = "Time", ylab = "Ridership", ylim = c(1300, 2300), bty = "l")
```


  


First3 <- window(ridership.ts, 1991, c(1994, 12))
plot(First3, xlab = "Time", ylab = "Ridership", ylim = c(1300, 2200), bty = "l" )
Last4 <- window(ridership.ts, 1997, c(2000, 12))
plot(Last4, xlab = "Time", ylab = "Ridership", ylim = c(1300, 2200), bty = "l" )


#Forecasts in the validation period from a quadratic trend model estimated from the training period


```{r}
nValid <- 36
nTrain <- length(ridership.ts) - nValid 
train.ts <- window(ridership.ts, start = c(1991, 1), end = c(1991, nTrain)) 
valid.ts <- window(ridership.ts, start = c(1991, nTrain + 1), end = c(1991, nTrain + nValid)) 
ridership.lm <- tslm(train.ts ~ trend + I(trend^2)) 
ridership.lm.pred <- forecast(ridership.lm, h = nValid, level = 0) 
plot(ridership.lm.pred, ylim = c(1300, 2600), ylab = "Ridership", xlab = "Time", bty = "l", xaxt = "n", xlim = c(1991,2006.25), main ="", flty = 2) 
axis(1, at = seq(1991, 2006, 1), labels = format(seq(1991, 2006, 1))) 
lines(ridership.lm$fitted, lwd = 2) 
lines(valid.ts)
```

#Computing predictive measures
```{r}
accuracy(ridership.lm.pred$mean, valid.ts)
```

#Calculating residuals

```{r}
names(ridership.lm.pred)
ridership.lm.pred$residuals
valid.ts - ridership.lm.pred$mean
```

#Histogram of residuals
```{r}
hist(ridership.lm.pred$residuals, ylab = "Frequency", xlab = "Forecast Error", bty = "l", main = "")
```

#computing the predictive measures for rollforward one-month-ahead forecasts in Table 3.3
 mean(abs(error)) sqrt(mean(error^2))


```{r}
fixed.nValid <- 36
fixed.nTrain <- length(ridership.ts) - fixed.nValid
stepsAhead <- 1
error <- rep(0, fixed.nValid - stepsAhead + 1)
percent.error <- rep(0, fixed.nValid - stepsAhead + 1)


for(j in fixed.nTrain:(fixed.nTrain + fixed.nValid - stepsAhead)) { 
  train.ts <- window(ridership.ts, start = c(1991, 1), end = c(1991, j)) 
  
  valid.ts <- window(ridership.ts, start = c(1991, j + stepsAhead), end = c(1991, j + stepsAhead)) 
  
  naive.pred <- naive(train.ts, h = stepsAhead) 
  error[j - fixed.nTrain + 1] <- valid.ts - naive.pred$mean[stepsAhead] 
  percent.error[j - fixed.nTrain + 1] <- error[j - fixed.nTrain + 1]/ valid.ts }
mean(abs(error)) 
sqrt(mean(error^2)) 
mean(abs(percent.error))
```

#Computation of naive and seasonal naive forecast and their predictive measures
```{r}
fixed.nValid <- 36
fixed.nTrain <- length(ridership.ts) - fixed.nValid
train.ts <- window(ridership.ts, start = c(1991,1), end = c(1991, fixed.nTrain))
valid.ts <- window(ridership.ts, start = c(1991, fixed.nTrain + 1), 
                   end = c(1991, fixed.nTrain + fixed.nValid))
naive.pred <- naive(train.ts, h = fixed.nValid)
snaive.pred <- snaive(train.ts, h = fixed.nValid)
accuracy(naive.pred, valid.ts)
accuracy(snaive.pred, valid.ts)

```

